\LARGE \bf Geometric Adaptive Control of Attitude Dynamics on $\SO$\\ with State Inequality Constraints

### 2015-09-25 ACC16.2.tex
moved proofs to appendix

\begin{prop}[Attitude Error Function] \label{prop:config_error}
We assume a desired attitude command \( ( R_d, \Omega_d = 0 ) \), the current attitude and angular velocity \( ( R, \Omega) \), and a state constraint~\cref{eqn:constraint} are given.
We define an attitude error function \( \Psi : \SO \times \SO \to \R \), and attitude error vector \( e_R : \SO \times \SO \to \R \), and an angular velocity error vector \( e_\Omega : \R^3 \to \R^3 \) as follows:
\begin{gather}
	\Psi(R) = A(R) B(R) \, , \label{eqn:psi} \\
	e_R = e_{R_A} B + A e_{R_B} \, , \label{eqn:eR} \\
	e_\Omega = \Omega \, , \label{eqn:eW}
\end{gather}
with
\begin{gather}
	A(R) = \frac{1}{2} \tr{G \left( I - R_d^T R\right)} \, , \label{eqn:A} \\
	B(R) = 1 - \frac{1}{\alpha} \ln \left( \frac{\cos \theta -  r^T R^T v}{1 + \cos \theta}\right) \, . \label{eqn:B} \\
	e_{R_A} = \frac{1}{2} \parenth{G R_d^T R - R^T R_d G} ^ \vee \, , \label{eqn:eRA} \\
	e_{R_B} = \frac{ \left( R^T v\right)^\vee r}{\alpha \left(r^T R^T v - \cos \theta \right)} \, . \label{eqn:eRB} 
\end{gather}	
where the matrix \( G \in \R^{3 \times 3} \) is defined as a diagonal matrix matrix for distinct, positive constants \( g_1, g_2, g_3 \in \R \).
Then, the following properties hold
\begin{enumerate}[(i)]
	\item \label{item:prop_psi_psd} \(\Psi\) is locally positive definite about \( R = R_d\)
	\item \label{item:prop_er} The variation of \( \Psi \) with respect to a variation of \( \delta R = R \hat{\eta} \) for \( \eta \in \R^3 \) is given by
	\begin{align}
		\dirDiff{\Psi}{R} &= \dirDiff{A}{R} \,  B(R) + A \,  \dirDiff{B}{R} \, . \\
	\end{align}
	\item \label{item:prop_era}The variation of \( A(R) \) with respect to a variation of \( \delta R = R \hat{\eta} \) for \( \eta \in \R^3 \) is given by
	\begin{align}
		\dirDiff{A}{R} &= \eta \cdot e_{R_A} \, .
	\end{align}
	\item \label{item:prop_erb} The variation of \( B(R) \) with respect to a variation of \( \delta R = R \hat{\eta} \) for \( \eta \in \R^3 \) is given by
	\begin{align}
		\dirDiff{B}{R} &= \eta \cdot e_{R_{B}} \, .
	\end{align}
	\item \label{item:prop_crit}The critical points of \( \Psi \),where \( e_R = 0\), are \( \braces{R_d} \cup \braces{R_d \exp(\pi \hat{s}} \) for \( s \in \braces{e_1, e_2, e_3} \) and \( R^T v = \pm r \).
	\item \label{item:prop_era_upbound}An upper bound of \( \norm{e_{R_A}} \) is given as:
	\begin{align}
		\norm{e_{R_A}}^2 \leq \frac{\Psi}{b_1} \, , \label{eqn:psi_lower_bound}
	\end{align}
	where the constant \( b_1 \) is given by \( b_1 = \frac{h_1}{h_2 + h_3} \) for 
	\begin{align*}
		h_1 &= \min{g_1 + g_2, g_2 + g_3 , g_3 + g_1} \, ,\\
		h_2 &= \min{\parenth{g_1 -g_2}^2,\parenth{g_2 -g_3}^2 , \parenth{g_3 -g_1}^2} \, ,\\
		h_3 &= \min{\parenth{g_1 + g_2}^2, \parenth{g_2 + g_3}^2 , \parenth{g_3 + g_1}^2} \, .				
	\end{align*}
\end{enumerate}
\end{prop}
\begin{proof}[Proof of~\Cref{prop:config_error}]\label{proof:config_error}
The proof of~\cref{item:prop_era_upbound} is available in~\cite{lee2011a}.
To prove~\cref{item:prop_psi_psd} we note that~\cref{eqn:A} is a locally positive definite function about \( R = R_d \)~\cite{bullo2004}.
The constraint angle is assumed \( \ang{0} \leq \theta \leq \ang{90} \) such that \( 0 \leq \cos \theta \).
The term \( r^T R^T v \) represents the cosine of the angle between the body fixed vector \( r \) and the inertial vector \( v \). 
It follows that
\begin{align*}
	0 \leq  \frac{\cos \theta -  r^T R^T v}{1 + \cos \theta} \leq 1 \, ,
\end{align*}
for all \( R \in \SO \). 
As a result, the negative logarithm is always positive and
\begin{align*}
	1 < B \, .
\end{align*}
The product~\cref{eqn:psi} is composed of two positive terms and is therefore also positive definite, and has a single critical point at \( R = R_d \).

To show~\cref{item:prop_crit}, we determine the critical points of~\cref{eqn:eRA}.
The attitude error between \( R \) and \( R_d \) is defined by \( Q \in \SO \) as
\begin{align*}
	Q = R_d^T R = 
	\begin{bmatrix}
		q_{11} & q_{12} & q_{13} \\
		q_{21} & q_{22} & q_{23} \\
		q_{31} & q_{32} & q_{33}
	\end{bmatrix} \, .
\end{align*}
Equating \( e_{R_{A}} = 0 \) yields the following equivalent relations
\begin{subequations}
\begin{align}
	g_3 q_{32} &= g_2 q_{23} \, , \\
	g_1 q_{13} &= g_3 q_{31} \, , \\
	g_2 q_{21} &= g_1 q_{12} \, .
\end{align}
\end{subequations}
These relations allow one to redefine the attitude error \( Q \) in terms of the independent upper diagonal elements.
Noting that \( Q \in \SO \) is an orthonormal matrix, with unit length row and columns gives
\begin{subequations}
\begin{align}
	\parenth{1 - \frac{g_1^2}{g_2^2}} q_{12}^2 + \parenth{1 - \frac{g_1^2}{g_3^2}}q_{13}^2 &= 0 \, , \\
	\parenth{1 - \frac{g_1^2}{g_2^2}} q_{12}^2 - \parenth{1 - \frac{g_2^2}{g_3^2}} q_{23}^2 &= 0 \, , \\
	\parenth{1-\frac{g_1^2}{g_3^2}} q_{13}^2 + \parenth{1 - \frac{g_2^2}{g_3^2}} q_{23}^2 &= 0 \, .
\end{align}
\end{subequations}
As a result, all the off diagonal terms of \( Q = R_d^T R \) are zero, namely \( q_{12}= q_{13} = q_{23} = 0 \).
Since \( Q \in \SO \), the diagonal elements \( q_{11}, q_{22}, \text{ and } q_{33} \) are either \( +1 \) or \(-1\).
There are four critical points of \( e_{R_A} \), the desired attitude \( R_d \) as well as rotations about each body fixed axis by \ang{180}.
The repulsive error vector \( e_{R_{B}} \) is zero only when the numerator \( \parenth{R^T v}^\wedge r = 0 \). 
This condition only occurs if the desired attitude results in the body fixed vector \( r \) becoming aligned with \(R^T v \) while simultaneously satisfying \(\braces{R_d} \cup \braces{R_d \exp(\pi \hat{s}} \) for \( s \in \braces{e_1, e_2, e_3} \) .
The desired equilibrium is \( e_R = 0 \) and \( A = 0\).
\end{proof}

\Cref{eqn:psi} is composed of an attractive term \( A (R) \) and a repulsive term \( B(R) \).
The attractive error function has been used for attitude control on \SO.
In order to incorporate the state inequality constraints we apply a logarithmic barrier function term.
Barrier functions are typically used in optimal control and motion planning applications.
As the boundary of the constraint is neared, or \( r^T R^T v \to \cos \theta \), the barrier term increases, \( B \to \infty\).
The logarithmic function is popular as it quickly decays away from the constraint boundary.
The positive constant \( \alpha \) serves to shape the barrier function.
As \( \alpha \) is increased the impact of \( B(R) \) is reduced away from the constraint boundary. 
In addition, it is possible to incorporate an arbitrary number of state inequality constraints of the form given in~\cref{eqn:constraint}.
The configuration error function may be reformulated as \( \Psi = A \bracket{1 + \sum B_i} \) where \( B_i = B - 1 \).
In this manner, one may enforce multiple state inequality constraints on the control system.
\begin{prop}[Error Dynamics]\label{prop:error_dyn}
	The attitude error dynamics for \( \Psi, e_R, e_\Omega \) satisfy 
	\begin{gather}
		\diff{}{t} \parenth{\Psi} = e_R \cdot e_\Omega \, , \label{eqn:psi_dot}\\
		\diff{}{t} \parenth{e_R} = \dot{e}_{R_A} B + e_{R_A} \dot{B} + \dot{A}e_{R_B} + A \dot{e}_{R_B} \, , \label{eqn:eR_dot}\\
		\diff{}{t} \parenth{e_{R_A}} = E(R, R_d) e_\Omega \, , \label{eqn:eRA_dot} \\
		\diff{}{t} \parenth{e_{R_B}} = F(R) e_\Omega \, , \label{eqn:eRB_dot} \\
		\diff{}{t} \parenth{A} = e_{R_A} \cdot e_\Omega \, , \label{eqn:A_dot} \\
		\diff{}{t} \parenth{B} = e_{R_B} \cdot e_\Omega \, , \label{eqn:B_dot} \\
		\diff{}{t} \parenth{e_\Omega} = J^{-1} \parenth{-\Omega \times J \Omega + u + W(R, \Omega) \Delta} \, , \label{eqn:eW_dot}
	\end{gather}
	where the matrices \(E(R,R_d), F(R) \in \R^{3\times3} \) are given by
	\begin{gather}
		E(R,R_d) = \frac{1}{2} \parenth{\tr{R^T R_d G}I - R^T R_d G} \, , \label{eqn:E} \\
		F(R) = \frac{1}{\alpha \parenth{r^T R^T v - \cos \theta}} \left[\parenth{v^T R r} I - R^T v r^T + \right. \nonumber \\
		\left. \frac{R^T \hat{v} R r v^T R \hat{r}}{\parenth{r^T R^T v - \cos \theta}}\right] \, . \label{eqn:F}
	\end{gather}
\end{prop}
\begin{proof}[Proof of~\Cref{prop:error_dyn}]\label{proof:error_dyn}
From the kinematics~\cref{eqn:Rdot} and noting that \( \dot{R}_d = 0 \) the time derivative of \( R_d^T R \) is given as
\begin{gather*}
	\diff{}{t} \parenth{R_d^T R} = R_d^T R \hat{e}_\Omega \, .
\end{gather*}
Applying this to the time derivative of~\cref{eqn:A} gives
\begin{gather*}
	\diff{}{t} (A) = -\frac{1}{2} \tr{G R_d^T R \hat{e}_\Omega} \, .
\end{gather*}
Applying~\cref{eqn:hat1} into this shows~\cref{eqn:A_dot}.
Next, the time derivative of the repulsive error function is given by
\begin{gather*}
	\diff{}{t} (B) = \frac{r_i^T \parenth{\hat{\Omega} R^T} v_i}{\alpha \parenth{r_i^T R^T v_i - \cos \theta}} \, .
\end{gather*}
Using the scalar triple product, given by~\cref{eqn:STP}, one can reduce this to~\cref{eqn:B_dot}.
The time derivative of the attractive attitude error vector, \( e_{R_A} \), is given by
\begin{gather*}
	\diff{}{t} ( e_{R_A}) = \frac{1}{2} \parenth{\hat{e}_\Omega R^T R_d G + (R^T R_d G)^T \hat{e}_\Omega}^\vee \, .
\end{gather*}
Using the hat map property given in~\cref{eqn:xAAx} this is further reduced to~\cref{eqn:eRA_dot,eqn:E}.

The time-derivative of the angular velocity error, \( e_\Omega \), is given by
\begin{align*}
	\diff{}{t} e_\Omega &= \dot{\Omega} \\
	&= J^{-1} \parenth{u - \Omega \times J \Omega + W(R,\Omega) \Delta } \, ,
\end{align*}
where we have used the dynamic equations of motion from~\cref{eqn:Wdot}. 
\end{proof}

@@

\begin{proof}[Proof of~\Cref{prop:att_control}]\label{proof:att_control}
	Consider the following Lyapunov function:
	\begin{gather}
		V = \frac{1}{2} e_\Omega \cdot J e_\Omega + k_R \Psi(R,R_d) \, . \label{eqn:v_nodist}
	\end{gather}
	From~\cref{item:prop_psi_psd}, we obtain
	\begin{gather*}
		V \geq 0 \, .
	\end{gather*}
	Using~\cref{eqn:eW_dot,eqn:psi_dot} with \( \Delta = 0 \), the time derivative of \( V \) is given by
	\begin{align}
		\dot{V} &= e_\Omega^T \parenth{ u - \Omega \times J \Omega } + k_R e_R \cdot e_\Omega \nonumber \\
		&= -k_\Omega \norm{e_\Omega}^2 \, . \label{eqn:vdot_nodist}
	\end{align}
	Since \( V \) is positive definite and \( \dot{V} \) is negative definite this implies that the zero equilibrium point \( e_R, e_\Omega \) is asymptotically stable.
\end{proof}

@@

\begin{proof}[Proof of~\Cref{prop:eR_dot_bound}]\label{proof:eR_dot_bound}
	We first show~\cref{eqn:A_bound,eqn:B_bound} through a proof by contradiction by noting that given the selected domain \( D \) we know the configuration error function is bounded by \( \Psi = A(R) B(R) < \psi\).
	If we assume that \( A > \psi \) this means that \( B < 1 \) in order to satisfy the domain \( \Psi < \psi \).
	However, this would be in contradiction with~\cref{proof:config_error} as \( B > 1 \), therefore \( A \) must always be less than \( \psi \).
	Next, we show that since the domain \( D \) does not contain the constraint region there is an upper bound of \(B\).
	We know that \( B \to \infty \) as the constraint is neared. 
	We assume that \( B > \psi \) which implies that \( A < 1 \).
	As \( \psi \) is increased this implies that the system is approaching the constraint and that \( A \to 0 \) in order to remain in the domain \( D \).
	However, since the chosen domain is assumed to exclude the constraint this situation is not possible and \( B < \psi \).
	This shows~\cref{eqn:A_bound,eqn:B_bound}.

	The bound on \( e_{R_A} \) is given in~\cref{eqn:psi_lower_bound} while \( e_{R_B} \) arises from the definition of the cross product \( \norm{a \times b} = \norm{a} \norm{b} \sin \theta \).
\end{proof}
\begin{prop}[Adaptive Attitude Control]\label{prop:adaptive_control}
Given  a desired attitude command \( (R_d, \Omega_d = 0 )\) and positive constants \( k_R, k_\Omega, k_\Delta, c \in \R \), we define a control input \( u \in \R^3\) and an adaptive update law for \( \bar{\Delta} \) as follows:
\begin{align}
	u &= - k_R e_R - k_\Omega e_\Omega + \Omega \times J \Omega - W \bar{\Delta} \, , \label{eqn:adaptive_control} \\
	\dot{\bar{\Delta}} &= k_\Delta W^T \parenth{e_\Omega + c e_R} \, . \label{eqn:delta_dot}
\end{align}
If \( c \) is chosen such that
\begin{gather}
	0 < c < \frac{4 k_R k_\Omega}{k_\Omega^2 + 4 k-R \lambda_M H} \, , \label{eqn:c_bound}
\end{gather}
  the zero equilibrium of the error vectors \( e_R, e_\Omega\) are asymptotically stable.
\end{prop}
\begin{proof}[Proof of~\cref{prop:adaptive_control}]\label{proof:adaptive_control}
	Consider the Lyapunov function \( V \) given by
	\begin{gather}
		V = \frac{1}{2} e_\Omega J e_\Omega + k_R \Psi + c J e_\Omega^T e_R + \frac{1}{2 k_\Delta} e_\Delta^T e_\Delta \, , \label{eqn:v_adapt}
	\end{gather}
	over the domain \( D \) in~\cref{eqn:domain}.
	From~\Cref{prop:eR_dot_bound}, the Lyapunov function is bounded in \( D \) by
	\begin{gather}
		V \leq z^T W z \, , \label{eqn:v_upper_bound}
	\end{gather}
	where \( e_\Delta = \Delta - \bar{\Delta} \), \( z = \begin{bmatrix}e_R & e_\Omega & e_\Delta\end{bmatrix} \) and the matrix \(W \in \R^{3 \times 3}\) is given by
	\begin{gather*}
		W = \begin{bmatrix}
			k_R \psi & \frac{1}{2} c \lambda_M & 0 \\
			\frac{1}{2} c \lambda_M & \frac{1}{2} \lambda_M & 0 \\
			0 & 0 & \frac{1}{2 k_\Delta}
		\end{bmatrix} \, .
	\end{gather*}
	The time derivative of \( V\) with the control inputs~\cref{eqn:adaptive_control} is defined as
	\begin{align}
		\dot{V} =& - k_\Omega e_\Omega^T e_\Omega + \parenth{e_\Omega + c e_R}^T W e_\Delta - k_R c e_R^T e_R \nonumber\\
		&- k_\Omega c e_R^T e_\Omega + c J e_\Omega^T \dot{e}_R - \frac{1}{k_\Delta} e_\Delta^T \dot{\bar{\Delta}} \, , \label{eqn:vdot}
	\end{align}
	where we used \( \dot{e}_\Delta = - \dot{\bar{\Delta}} \).
	The terms linearly dependent on \( e_\Delta\) are combined with~\cref{eqn:delta_dot} to yield
	\begin{align*}
		 e_\Delta^T \parenth{W^T \parenth{e_\Omega + c e_R} - \frac{1}{k_\Delta} \dot{\bar{\Delta}}} = 0 \, . 
	\end{align*}
	Using~\Cref{prop:eR_dot_bound} an upper bound on \( \dot{V} \) is written as
	\begin{gather*}
		\dot{V} \leq -z^T M z \, ,
	\end{gather*}
	where the matrix \( M \in \R^{2 \times 2} \) is 
	\begin{gather}
		M = \begin{bmatrix}
			k_R c & \frac{k_\Omega c}{2} \\
			\frac{k_\Omega c}{2} & k_\Omega - c \lambda_M H
		\end{bmatrix} \, .
	\end{gather}
	If \( c \) is chosen such that~\cref{eqn:c_bound} is satisfied the matrix \( M \) is positive definite.
	This implies that the zero equilibrium \( e_R, e_\Omega\) is asymptotically stable. 
	The domain \( D \) does that include the undesired equilibria and ensures that as \( e_R \to 0 \) the attitude also converges \( R \to R_d \).
\end{proof}

@@

\section{Numerical Examples}
We demonstrate the performance of the proposed control system via numerical simulation.
The inertia tensor of a rigid body is given as
\begin{gather*}
	J = \begin{bmatrix}
	\num{5.57e-3} & \num{6.17e-5} & \num{-2.50e-5} \\
	\num{6.17e-5} & \num{5.57e-3} & \num{1.00e-5} \\
	\num{-2.50e-5} & \num{1.00e-5} & \num{1.05e-2}
	\end{bmatrix} \si{\kilo\gram\meter\squared} \, .
\end{gather*} 
The control system parameters are chosen as
\begin{gather*}
	G = \text{diag} \parenth{0.9,1.1,1.0}\, , \quad k_R = 0.4 \, , \quad	k_\Omega = 0.296 \, ,\\
	c = 1.0 \, , \quad k_\Delta = 0.5 \, , \quad \alpha = 15 \, .
\end{gather*}
A body fixed sensor is defined as
\begin{gather*}
	r = \begin{bmatrix} 1 & 0 & 0 \end{bmatrix} \, ,
\end{gather*}
while multiple inequality constraints are defined in~\Cref{tab:constraints}.
The simulation parameters are chosen to emulate those found in~\cite{lee2011b}.
\begin{table}[htbp]
\begin{center}\begin{tabular}{lc}
Constraint Vector & Angle \\ \hline \hline 
\(\begin{bmatrix} 0.174 &-0.934 & -0.034 \end{bmatrix}\) & \ang{40} \\ \hline 
\(\begin{bmatrix} 0 & 0.7071 & 0.7071 \end{bmatrix}\) & \ang{40} \\ \hline 
\(\begin{bmatrix} -0.853 & 0.436 & -0.286 \end{bmatrix}\) & \ang{40} \\ \hline 
\(\begin{bmatrix} -0.122 -0.140 -0.983 \end{bmatrix}\) & \ang{20}\end{tabular} 
\caption{Simulation Parameters~\label{tab:constraints}}
\end{center}
\end{table}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix

>>>>

\begin{prop}[Attitude Error Function] \label{prop:config_error}
We assume a desired attitude command \( ( R_d, \Omega_d = 0 ) \), the current attitude and angular velocity \( ( R, \Omega) \), and a state constraint~\cref{eqn:constraint} are given.
We define an attitude error function \( \Psi : \SO \times \SO \to \R \), and attitude error vector \( e_R : \SO \times \SO \to \R \), and an angular velocity error vector \( e_\Omega : \R^3 \to \R^3 \) as follows:
\begin{gather}
	\Psi(R) = A(R) B(R) \, , \label{eqn:psi} \\
	e_R = e_{R_A} B + A e_{R_B} \, , \label{eqn:eR} \\
	e_\Omega = \Omega \, , \label{eqn:eW}
\end{gather}
with
\begin{gather}
	A(R) = \frac{1}{2} \tr{G \left( I - R_d^T R\right)} \, , \label{eqn:A} \\
	B(R) = 1 - \frac{1}{\alpha} \ln \left( \frac{\cos \theta -  r^T R^T v}{1 + \cos \theta}\right) \, . \label{eqn:B} \\
	e_{R_A} = \frac{1}{2} \parenth{G R_d^T R - R^T R_d G} ^ \vee \, , \label{eqn:eRA} \\
	e_{R_B} = \frac{ \left( R^T v\right)^\vee r}{\alpha \left(r^T R^T v - \cos \theta \right)} \, . \label{eqn:eRB} 
\end{gather}	
where the matrix \( G \in \R^{3 \times 3} \) is defined as a diagonal matrix matrix for distinct, positive constants \( g_1, g_2, g_3 \in \R \).
Then, the following properties hold
\begin{enumerate}[(i)]
	\item \label{item:prop_psi_psd} \(\Psi\) is locally positive definite about \( R = R_d\)
	\item \label{item:prop_er} The variation of \( \Psi \) with respect to a variation of \( \delta R = R \hat{\eta} \) for \( \eta \in \R^3 \) is given by
	\begin{align}
		\dirDiff{\Psi}{R} &= \dirDiff{A}{R} \,  B(R) + A \,  \dirDiff{B}{R} \, . \\
	\end{align}
	\item \label{item:prop_era}The variation of \( A(R) \) with respect to a variation of \( \delta R = R \hat{\eta} \) for \( \eta \in \R^3 \) is given by
	\begin{align}
		\dirDiff{A}{R} &= \eta \cdot e_{R_A} \, .
	\end{align}
	\item \label{item:prop_erb} The variation of \( B(R) \) with respect to a variation of \( \delta R = R \hat{\eta} \) for \( \eta \in \R^3 \) is given by
	\begin{align}
		\dirDiff{B}{R} &= \eta \cdot e_{R_{B}} \, .
	\end{align}
	\item \label{item:prop_crit}The critical points of \( \Psi \),where \( e_R = 0\), are \( \braces{R_d} \cup \braces{R_d \exp(\pi \hat{s}} \) for \( s \in \braces{e_1, e_2, e_3} \) and \( R^T v = \pm r \).
	\item \label{item:prop_era_upbound}An upper bound of \( \norm{e_{R_A}} \) is given as:
	\begin{align}
		\norm{e_{R_A}}^2 \leq \frac{\Psi}{b_1} \, , \label{eqn:psi_lower_bound}
	\end{align}
	where the constant \( b_1 \) is given by \( b_1 = \frac{h_1}{h_2 + h_3} \) for 
	\begin{align*}
		h_1 &= \min{g_1 + g_2, g_2 + g_3 , g_3 + g_1} \, ,\\
		h_2 &= \min{\parenth{g_1 -g_2}^2,\parenth{g_2 -g_3}^2 , \parenth{g_3 -g_1}^2} \, ,\\
		h_3 &= \min{\parenth{g_1 + g_2}^2, \parenth{g_2 + g_3}^2 , \parenth{g_3 + g_1}^2} \, .				
	\end{align*}
\end{enumerate}
\end{prop}
\begin{proof}
See~\Cref{proof:config_error}
\end{proof}

\Cref{eqn:psi} is composed of an attractive term \( A (R) \) and a repulsive term \( B(R) \).
The attractive error function has been used for attitude control on \SO.
In order to incorporate the state inequality constraints we apply a logarithmic barrier function term.
Barrier functions are typically used in optimal control and motion planning applications.
As the boundary of the constraint is neared, or \( r^T R^T v \to \cos \theta \), the barrier term increases, \( B \to \infty\).
The logarithmic function is popular as it quickly decays away from the constraint boundary.
The positive constant \( \alpha \) serves to shape the barrier function.
As \( \alpha \) is increased the impact of \( B(R) \) is reduced away from the constraint boundary. 
In addition, it is possible to incorporate an arbitrary number of state inequality constraints of the form given in~\cref{eqn:constraint}.
The configuration error function may be reformulated as \( \Psi = A \bracket{1 + \sum B_i} \) where \( B_i = B - 1 \).
In this manner, one may enforce multiple state inequality constraints on the control system.
\begin{prop}[Error Dynamics]\label{prop:error_dyn}
	The attitude error dynamics for \( \Psi, e_R, e_\Omega \) satisfy 
	\begin{gather}
		\diff{}{t} \parenth{\Psi} = e_R \cdot e_\Omega \, , \label{eqn:psi_dot}\\
		\diff{}{t} \parenth{e_R} = \dot{e}_{R_A} B + e_{R_A} \dot{B} + \dot{A}e_{R_B} + A \dot{e}_{R_B} \, , \label{eqn:eR_dot}\\
		\diff{}{t} \parenth{e_{R_A}} = E(R, R_d) e_\Omega \, , \label{eqn:eRA_dot} \\
		\diff{}{t} \parenth{e_{R_B}} = F(R) e_\Omega \, , \label{eqn:eRB_dot} \\
		\diff{}{t} \parenth{A} = e_{R_A} \cdot e_\Omega \, , \label{eqn:A_dot} \\
		\diff{}{t} \parenth{B} = e_{R_B} \cdot e_\Omega \, , \label{eqn:B_dot} \\
		\diff{}{t} \parenth{e_\Omega} = J^{-1} \parenth{-\Omega \times J \Omega + u + W(R, \Omega) \Delta} \, , \label{eqn:eW_dot}
	\end{gather}
	where the matrices \(E(R,R_d), F(R) \in \R^{3\times3} \) are given by
	\begin{gather}
		E(R,R_d) = \frac{1}{2} \parenth{\tr{R^T R_d G}I - R^T R_d G} \, , \label{eqn:E} \\
		F(R) = \frac{1}{\alpha \parenth{r^T R^T v - \cos \theta}} \left[\parenth{v^T R r} I - R^T v r^T + \right. \nonumber \\
		\left. \frac{R^T \hat{v} R r v^T R \hat{r}}{\parenth{r^T R^T v - \cos \theta}}\right] \, . \label{eqn:F}
	\end{gather}
\end{prop}
\begin{proof}
See~\Cref{proof:error_dyn}
\end{proof}

@@

\begin{proof}
See~\Cref{proof:att_control}
\end{proof}

@@

\begin{proof}
See~\Cref{proof:eR_dot_bound}
\end{proof}
\begin{prop}[Adaptive Attitude Control]\label{prop:adaptive_control}
Given  a desired attitude command \( (R_d, \Omega_d = 0 )\) and positive constants \( k_R, k_\Omega, k_\Delta, c \in \R \), we define a control input \( u \in \R^3\) and an adaptive update law for \( \bar{\Delta} \) as follows:
\begin{align}
	u &= - k_R e_R - k_\Omega e_\Omega + \Omega \times J \Omega - W \bar{\Delta} \, , \label{eqn:adaptive_control} \\
	\dot{\bar{\Delta}} &= k_\Delta W^T \parenth{e_\Omega + c e_R} \, . \label{eqn:delta_dot}
\end{align}
If \( c \) is chosen such that
\begin{gather}
	0 < c < \frac{4 k_R k_\Omega}{k_\Omega^2 + 4 k-R \lambda_M H} \, , \label{eqn:c_bound}
\end{gather}
  the zero equilibrium of the error vectors \( e_R, e_\Omega\) are asymptotically stable.
\end{prop}
\begin{proof}
See~\Cref{proof:adaptive_control}
\end{proof}

@@

\section{Numerical Examples}
We demonstrate the performance of the proposed control system via numerical simulation.
The inertia tensor of a rigid body is given as
\begin{gather*}
	J = \begin{bmatrix}
	\num{5.57e-3} & \num{6.17e-5} & \num{-2.50e-5} \\
	\num{6.17e-5} & \num{5.57e-3} & \num{1.00e-5} \\
	\num{-2.50e-5} & \num{1.00e-5} & \num{1.05e-2}
	\end{bmatrix} \si{\kilo\gram\meter\squared} \, .
\end{gather*} 
The control system parameters are chosen as
\begin{gather*}
	G = \text{diag} \parenth{0.9,1.1,1.0}\, , \quad k_R = 0.4 \, , \quad	k_\Omega = 0.296 \, ,\\
	c = 1.0 \, , \quad k_\Delta = 0.5 \, , \quad \alpha = 15 \, .
\end{gather*}
A body fixed sensor is defined as
\begin{gather*}
	r = \begin{bmatrix} 1 & 0 & 0 \end{bmatrix} \, ,
\end{gather*}
while multiple inequality constraints are defined in~\Cref{tab:constraints}.
The simulation parameters are chosen to emulate those found in~\cite{lee2011b}.
\begin{table}[htbp]
\caption{Simulation Parameters~\label{tab:constraints}}
\begin{center}\begin{tabular}{lc}
Constraint Vector & Angle \\ \hline \hline 
\(\begin{bmatrix} 0.174 &-0.934 & -0.034 \end{bmatrix}\) & \ang{40} \\ \hline 
\(\begin{bmatrix} 0 & 0.7071 & 0.7071 \end{bmatrix}\) & \ang{40} \\ \hline 
\(\begin{bmatrix} -0.853 & 0.436 & -0.286 \end{bmatrix}\) & \ang{40} \\ \hline 
\(\begin{bmatrix} -0.122 -0.140 -0.983 \end{bmatrix}\) & \ang{20}\end{tabular} 
\end{center}
\end{table}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix
\subsection{Proof of~\Cref{prop:config_error}}\label{proof:config_error}
The proof of~\cref{item:prop_era_upbound} is available in~\cite{lee2011a}.
To prove~\cref{item:prop_psi_psd} we note that~\cref{eqn:A} is a locally positive definite function about \( R = R_d \)~\cite{bullo2004}.
The constraint angle is assumed \( \ang{0} \leq \theta \leq \ang{90} \) such that \( 0 \leq \cos \theta \).
The term \( r^T R^T v \) represents the cosine of the angle between the body fixed vector \( r \) and the inertial vector \( v \). 
It follows that
\begin{align*}
	0 \leq  \frac{\cos \theta -  r^T R^T v}{1 + \cos \theta} \leq 1 \, ,
\end{align*}
for all \( R \in \SO \). 
As a result, the negative logarithm is always positive and
\begin{align*}
	1 < B \, .
\end{align*}
The product~\cref{eqn:psi} is composed of two positive terms and is therefore also positive definite, and has a single critical point at \( R = R_d \).

To show~\cref{item:prop_crit}, we determine the critical points of~\cref{eqn:eRA}.
The attitude error between \( R \) and \( R_d \) is defined by \( Q \in \SO \) as
\begin{align*}
	Q = R_d^T R = 
	\begin{bmatrix}
		q_{11} & q_{12} & q_{13} \\
		q_{21} & q_{22} & q_{23} \\
		q_{31} & q_{32} & q_{33}
	\end{bmatrix} \, .
\end{align*}
Equating \( e_{R_{A}} = 0 \) yields the following equivalent relations
\begin{subequations}
\begin{align}
	g_3 q_{32} &= g_2 q_{23} \, , \\
	g_1 q_{13} &= g_3 q_{31} \, , \\
	g_2 q_{21} &= g_1 q_{12} \, .
\end{align}
\end{subequations}
These relations allow one to redefine the attitude error \( Q \) in terms of the independent upper diagonal elements.
Noting that \( Q \in \SO \) is an orthonormal matrix, with unit length row and columns gives
\begin{subequations}
\begin{align}
	\parenth{1 - \frac{g_1^2}{g_2^2}} q_{12}^2 + \parenth{1 - \frac{g_1^2}{g_3^2}}q_{13}^2 &= 0 \, , \\
	\parenth{1 - \frac{g_1^2}{g_2^2}} q_{12}^2 - \parenth{1 - \frac{g_2^2}{g_3^2}} q_{23}^2 &= 0 \, , \\
	\parenth{1-\frac{g_1^2}{g_3^2}} q_{13}^2 + \parenth{1 - \frac{g_2^2}{g_3^2}} q_{23}^2 &= 0 \, .
\end{align}
\end{subequations}
As a result, all the off diagonal terms of \( Q = R_d^T R \) are zero, namely \( q_{12}= q_{13} = q_{23} = 0 \).
Since \( Q \in \SO \), the diagonal elements \( q_{11}, q_{22}, \text{ and } q_{33} \) are either \( +1 \) or \(-1\).
There are four critical points of \( e_{R_A} \), the desired attitude \( R_d \) as well as rotations about each body fixed axis by \ang{180}.
The repulsive error vector \( e_{R_{B}} \) is zero only when the numerator \( \parenth{R^T v}^\wedge r = 0 \). 
This condition only occurs if the desired attitude results in the body fixed vector \( r \) becoming aligned with \(R^T v \) while simultaneously satisfying \(\braces{R_d} \cup \braces{R_d \exp(\pi \hat{s}} \) for \( s \in \braces{e_1, e_2, e_3} \) .
The desired equilibrium is \( e_R = 0 \) and \( A = 0\).

\subsection{Proof of~\Cref{prop:error_dyn}}\label{proof:error_dyn}
From the kinematics~\cref{eqn:Rdot} and noting that \( \dot{R}_d = 0 \) the time derivative of \( R_d^T R \) is given as
\begin{gather*}
	\diff{}{t} \parenth{R_d^T R} = R_d^T R \hat{e}_\Omega \, .
\end{gather*}
Applying this to the time derivative of~\cref{eqn:A} gives
\begin{gather*}
	\diff{}{t} (A) = -\frac{1}{2} \tr{G R_d^T R \hat{e}_\Omega} \, .
\end{gather*}
Applying~\cref{eqn:hat1} into this shows~\cref{eqn:A_dot}.
Next, the time derivative of the repulsive error function is given by
\begin{gather*}
	\diff{}{t} (B) = \frac{r_i^T \parenth{\hat{\Omega} R^T} v_i}{\alpha \parenth{r_i^T R^T v_i - \cos \theta}} \, .
\end{gather*}
Using the scalar triple product, given by~\cref{eqn:STP}, one can reduce this to~\cref{eqn:B_dot}.
The time derivative of the attractive attitude error vector, \( e_{R_A} \), is given by
\begin{gather*}
	\diff{}{t} ( e_{R_A}) = \frac{1}{2} \parenth{\hat{e}_\Omega R^T R_d G + (R^T R_d G)^T \hat{e}_\Omega}^\vee \, .
\end{gather*}
Using the hat map property given in~\cref{eqn:xAAx} this is further reduced to~\cref{eqn:eRA_dot,eqn:E}.
